Detecting Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

Email spam and techniques with which to combat it has been studied 
extensively. Like
This paper examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

@@

We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
over 1.3 million active members. This data includes a corpus of over a 
100 million private messages and another 2 million messages that have been 
labeled as spam by users. Other data includes 40 million or so "wall" 
comments, 5 million photos, and 8 million photo comments.

@@

We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
message text, we used the following features:
\begin{itemize}
\item Sex of sender
\item Age of sender
\item Age of recipient
\item Account lifetime
\item Month of birth 
\item Whether sender has friends (boolean)
\item Whether sender has photo(s) (boolean)
\item Profile country and IP-detected country match (boolean)
\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) (boolean)
\end{itemize}

@@

% Modified from Reynold's HW1  write-up
%We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
%Then, a document from class \(c\) is generated.
%We tried a multinomial model and a Bernoulli model to model this process.
%\begin{itemize}
%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
%\end{itemize}
%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
%
%

@@

The decline in accuracy after adding expert features is most likely explained by the "naive" assumption in this 
algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
these features during classification. The wealth of expert features available to a SNS, many of which are 
not independent, makes this the least compelling classifier choice that we examined.

@@

We tried a number of values for $\lambda$: 1, 10, 50, 100, 500, 1000, 1500, 3000. Figure \ref{fig:lambda}

@@

% LR + expert
\begin{table}
\begin{tabular}{l|l||l|l}
\multicolumn{2}{c}{\textbf{Ham}} & 
\multicolumn{2}{c}{\textbf{Spam}} \\
\hline
\textbf{Feature} & \textbf{Weight} & \textbf{Feature} & \textbf{Weight} \\
\hline

@@

\end{tabular}
\caption{Linear Regression top features (bag-of-words + expert features) with weights. Expert features
are denoted by the 'F-' prefix.}
\label{tab:lrfeats}
\end{table}

@@

Our initial results using the Pegasos SVM solver included in ScalaNLP were disappointing. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
"2012," "wish," and "you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of January.

\begin{table}
\begin{tabular}{l|l||l|l}
\multicolumn{2}{c}{\textbf{Bag-of-Words}} & 
\multicolumn{2}{c}{\textbf{Expert Features}} \\
\hline
\textbf{Feature} & \textbf{Ham Weight} & \textbf{Feature} & \textbf{Ham Weight} \\
\hline

year & 0.2253 & F-account\_lifetime & -0.98\\
new & 0.225 & F-friends & -0.3671\\
am & -0.1722 & F-recip\_age & -0.2129\\
happy & 0.16 & F-age & -0.183\\
d & 0.1513 & i & -0.05231\\
ok & -0.1497 & you & -0.0483\\
yahoo & -0.136 & to & -0.04026\\
it & 0.12 & and & -0.03636\\
me & -0.118 & F-month\_of\_birth & -0.03452\\
too & 0.1173 & F-country\_match & -0.0269\\
i'm & 0.1116 & me & -0.0224\\
haha & 0.0998 & am & -0.022\\
can & -0.0971 & my & -0.0209\\
okay & -0.0947 & ? & 0.0199\\
dear & -0.0882 & ? & 0.0191\\
p & 0.0862 & a & -0.01573\\
here & -0.0844 & your & -0.01554\\
2012 & 0.084 & F-photo\_exists & -0.0146\\
wish & 0.0836 & the & -0.01412\\
doing & -0.0779 & f-sex & -0.01258\\

@@

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:drop}
figure \ref{fig:friends}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:roc-25}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/roc-25.pdf}
    \caption{LIBSVM ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features on 50k messages.}
    \label{fig:roc-25}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/sex.pdf}
    \caption{Sender sex as stated on profile.}
    \label{fig:sendsex}
\end{figure}

>>>>

Email spam and techniques with which to combat it has been studied 
extensively. Like
This paper examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
``419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

@@

We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
over 1.3 million active members. This data includes a corpus of over a 
100 million private messages and another 2 million messages that have been 
labeled as spam by users. Other data includes 40 million or so ``wall" 
comments, 5 million photos, and 8 million photo comments.

@@

We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
message text, we used the following features:
\begin{itemize}
\item Sex of sender
\item Age of sender
\item Age of recipient
\item Account lifetime
\item Month of birth 
\item Whether sender has friends , boolean
\item Whether sender has photo(s) , boolean
\item Profile country and IP-detected country match , boolean
\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) , boolean
\end{itemize}

@@

% Modified from Reynold's HW1  write-up
%We use a na\``ive Bayes model, in which the documents are generated according to the following process.  
%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
%Then, a document from class \(c\) is generated.
%We tried a multinomial model and a Bernoulli model to model this process.
%\begin{itemize}
%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
%\end{itemize}
%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
%
%

@@

The decline in accuracy after adding expert features is most likely explained by the ``naive" assumption in this 
algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
these features during classification. The wealth of expert features available to a SNS, many of which are 
not independent, makes this the least compelling classifier choice that we examined.

@@

We tried a number of values for $\lambda$: 1, 10, 50, 100, 500, 1000, 1500, 3000. Figure \ref{fig:lambda}
\ref{fig:roclin}

@@

% LR + expert
\begin{table}
\begin{tabular}{l|r||l|r}
\multicolumn{2}{c}{\textbf{Ham}} & 
\multicolumn{2}{c}{\textbf{Spam}} \\
\hline
\textbf{Feature} & \textbf{Weight} & \textbf{Feature} & \textbf{Weight} \\
\hline

@@

\end{tabular}
\caption{Linear Regression top features (bag-of-words \& expert features) with weights. Expert features
are denoted by an 'F-' prefix.}
\label{tab:lrfeats}
\end{table}

@@

Our initial results using the Pegasos SVM solver included in ScalaNLP were lower than expected. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words ``year," ``new," ``happy,"  
``2012," ``wish," and ``you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of 2012. The table shows that three expert features: account lifetime, 
friends and recipient age, are weighted very heavily in comparison to the remaining features. These 
skewed weights (and the inability to tune the SVM solver via ScalaNLP) prompted us to test LIBSVM 
on the same data.

\begin{table}
\begin{tabular}{l|r||l|r}
\multicolumn{2}{c}{\textbf{Bag-of-Words}} & 
\multicolumn{2}{c}{\textbf{Expert Features}} \\
\hline
\textbf{Feature} & \textbf{Ham Wt.} & \textbf{Feature} & \textbf{Ham Wt.} \\
\hline

year & 0.2253 & F-account\_lifetime & -0.9800\\
new & 0.2250 & F-friends & -0.3671\\
am & -0.1722 & F-recip\_age & -0.2129\\
happy & 0.1600 & F-age & -0.1830\\
d & 0.1513 & i & -0.05231\\
ok & -0.1497 & you & -0.0483\\
yahoo & -0.1360 & to & -0.04026\\
it & 0.1200 & and & -0.03636\\
me & -0.1180 & F-month\_of\_birth & -0.03452\\
too & 0.1173 & F-country\_match & -0.0269\\
i'm & 0.1116 & me & -0.0224\\
haha & 0.0998 & am & -0.0220\\
can & -0.0971 & my & -0.0209\\
okay & -0.0947 & ? & 0.0199\\
dear & -0.0882 & ? & 0.0191\\
p & 0.0862 & a & -0.01573\\
here & -0.0844 & your & -0.01554\\
2012 & 0.084 & F-photo\_exists & -0.0146\\
wish & 0.0836 & the & -0.01412\\
doing & -0.0779 & f-sex & -0.01258\\

@@

Our results with LIBSVM's RBF kernel were consistently better than any of the other classifiers we tested. We 
evaluated the classifier's performance on five pairs of bag-of-words and expert feature data sets ranging in 
size from 2000 to 100k messages. Table \ref{tab:svmacc} shows results of 10-fold cross-validation. Our mean 
accuracy over the bag-of-words data sets was 85.73\% and 98.41\% over the bag-of-words with expert features. We 
observed a slight decrease in performance on the bag-of-words data sets as the data size grew. Figure \ref{fig:roc-25}
shows the AUCs and ROC curves for the 50k data sets, making clear the the significantly greater AUC resulting 
from the addition of expert features.

\begin{table}
\begin{tabular}{r|r|r}
\hline
\textbf{Dataset Size} & \textbf{Bag-of-Words} & \textbf{Expert Features} \\
\hline
2000 & 86.70\% & 98.35\% \\
10000 & 86.23\% & 98.49\% \\
22000 & 86.81\% & 98.33\% \\
50000 & 84.59\% & 98.43\% \\
100000 & 84.33\% & 98.45\% \\
\end{tabular}
\caption{SVM 10-fold cross-validation accuracy over data sets of varying sizes (half spam, half ham)} 
\label{tab:svmacc}
\end{table}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/roc-25.pdf}
    \caption{LIBSVM ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features on 50k messages 
     (25k spam, 25k ham)}
    \label{fig:roc-25}
\end{figure}

@@

%\begin{figure}[h]
%    \centering
%    \includegraphics[width=\linewidth]{figures/sex.pdf}
%    \caption{Sender sex as stated on profile.}
%    \label{fig:sendsex}
%\end{figure}
%
